\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-03-11 paper/hubble_paperclip.tex
Add overview fig

\newcommand{\githubmaster}{\href{https://github.com/smsharma/HubbleCLIP}{\faGithub}\xspace}

@@

\lstdefinestyle{mystyle}{ 
  backgroundcolor=\color{backcolour}, 
  commentstyle=\color{codegreen},
   keywordstyle=\color{codegreen}, 
   numberstyle=\tiny\color{codegray}, 
   stringstyle=\color{codegreen},
   basicstyle=\ttfamily\footnotesize\linespread{1.1},
   breakatwhitespace=false, 
   breaklines=true, captionpos=b, 
   keepspaces=true, numbers=left, 
   numbersep=8pt, showspaces=false, 
   showstringspaces=false, showtabs=false, 
   frame=single, 
   framerule=0.2pt, 
   rulecolor=\color{codegray}, 
   tabsize=2, 
   aboveskip=1.5ex, 
   belowskip=1.5ex,
   xleftmargin=15pt, 
   xrightmargin=15pt, 
   extendedchars=true, 
   columns=flexible, 
   linewidth=\textwidth 
  }

@@

Machine learning (ML) is starting to have a significant impact in the sciences, with astrophysics being no exception.
%
ML methods have demonstrated promise at every stage of the research pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
%
Until recently, most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
%
The {Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
%
% The paradigm has been highly successful in domains like computer vision and natural language processing, as demonstrated by the widespread use and adoption of tools like ChatGPT, Dall-E, Stable Diffusion, and CLIP.
%
% At the heart of the paradigm lies the triumph of scale -- scaling up model size, dataset size, and compute.
% %
These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream in different ways.
%
Foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.

@@

In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations. 
%
Concretely, here we showcase the method using observations imaged by the \hubble Space Telescope (HST).
%
We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
%
Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
%

@@

% \SM{Put in this work we at the start of the paragraph. Also put the paragraph right after AstroLlama, reworking some of the logic.} \SM{Define PAPERCLIP, funny footnote about PAPERCLIFT.}

@@

@@

% \section{Dataset Construction}
% \label{sec:dataset}

% We curate a dataset of \hubble Space Telescope (HST) observations and corresponding text descriptions from publicly available sources. The dataset consists of images from the HST and associated proposal abstracts from the Hubble Space Telescope Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}}, which contains all accepted HST proposals since the telescope's launch on April 24, 1990. Our dataset covers observations up to Cycle 30, which began collecting data in 2022.

% To generate captions for the observations, we use the proposal abstracts directly or optionally summarize them using guided language model generation (described in Sec.\ref{sec:summarization}).

% The images in this dataset exhibit specific characteristics and artifacts inherent to the HST's data acquisition and processing pipeline, setting them apart from typical natural image datasets used for large-scale pre-training of foundation models \citep{deng2009imagenet}. This difference highlights the necessity of fine-tuning models on domain-specific data.

% We emphasize that all data used in this work is publicly available, ensuring transparency and enabling reproducibility. The HST images can be accessed through the Hubble Legacy Archive\footnote{\url{https://hla.stsci.edu/}}, while the proposal abstracts are available in the Hubble Space Telescope Proposal Abstracts Catalog.

@@

The guided generation process ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
% \SM{Have it be centered}
% \begin{lstlisting}[language=Python]
% {
%   'objects_and_phenomena': ['star forming galaxy', 'lensed galaxy', ...], 
%   'science_use_cases': ['measure lensing magnification', 'probe spectral energy distributions', ...]
% }
% \end{lstlisting}
\begin{center}
  \begin{jsoncode}
    \centering
          \color{black}\{
            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}], 
            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}]
          \color{black}\}
  \end{jsoncode}
  \end{center}
which is then used to construct the summarized caption by combining the two key elements.
%
Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
%
% We emphasize that the intended goal of summarization-via-guided-generation is to increase the signal between text and images by standardizing the captions used for fine-tuning the CLIP model, and compare the quantitative performance of the model vs using the raw abstracts in Sec.~\ref{sec:results}.
%
We train separate models using the raw abstracts and the LLM-generated summaries, and compare their performance on downstream tasks in Sec.~\ref{sec:results}.
%
We note that, even after summarization, the association signal is expected to be noisy, since parts of the summarized caption may not be directly descriptive of the observed images. The goal of the fine-tuning process is to leverage the signal contained in this noisy association.

% Finally, in order to test a further compression of the proposal abstracts, we use LLM-guided generation to also produce a list of single-concept summaries (e.g., `irregular galaxy', `quasars', `Galactic bulge', \ldots), and test whether this can lead to meaningful, generalizable learned associations with observations.
% %
% The prompts and schemata for generating these discrete categories and assigning observations to them are described in Apps.~\ref{app:singleconcept} and \ref{app:singleconceptassignments} respectively. \SM{Each abstract is assigned one summary. The generation of the single concepts is different }

% \begin{landscape}
%   \begin{table}[h!]
%     \renewcommand{\arraystretch}{2}
%       \centering
%       \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6.5cm}}
%           \toprule
%           \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
%           \cmidrule(r){3-4}
%           & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
%           \midrule
%           \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
%           \midrule
%           \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
%           \midrule
%           \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
%           \midrule
%           \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
%           \bottomrule
%       \end{tabular}
%       \caption{Examples of the clipped \hubble proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (right-most two columns), separately extracting objects and phenomena as well as potential downstream science use cases.}
%       \label{tab:datasetsumm}
%   \end{table}
%   \end{landscape}

@@

% Similarly, Fig.~\ref{fig:sim_valtrain} (top right) shows the distribution of cosine similarities between corresponding image-text pairs for the single-concept captions (solid red) and the shuffled versions (dashed red).
% %
% The distribution for corresponding pairs is only modestly shifted compared to the shuffled version, indicating that assigning $\mathcal O(400)$ labels to the images using an LLM results in a weaker association signal compared to using noisy captions.
% %
% The purple line shows the distribution of cosine similarities between corresponding pairs with single-concept summaries, but evaluated using a model trained on the LLM-summarized abstracts.
% %
% This distribution is more skewed towards higher similarity values, indicating that the model trained on diverse caption can still better capture the association signal in the single-concept summaries compared to one trained specifically using the single-concept summaries.

@@

The model fine-tuned with summarized abstracts, meanwhile, shows strikingly different behavior (Tab.~\ref{tab:tti}).
%
% For example, it is able to return images with processing and assembly artifacts particular to HST (e.g., the lines through the middle in some images), which typically receive low similarity scores when using the base model.
%
The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
%
Images of Jupiter are returned for the \texttt{Jupiter} query, with the planet clearly visible in all four images.
%
Supernova SN1987 itself can be seen in the three closest images for the \texttt{SN1987A} query with the fourth image being a supernova remnant.
%
Cluster-scale as well as galaxy-scale gravitational lenses are returned by the \texttt{strong lensing} query, with lensing patterns visible in the images.

@@

We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., objects and use cases) from a curated list as described in Sec.~\ref{sec:eval}.
%
We show the result of image-to-text retrieval in Tab.~\ref{tab:itt}, for the base (second column) as well as summary fine-tuned (third column) models, using four observations (left-most column) from the validation set.
% %
% We curate a list of possible text associations by querying the \textsc{Claude}\footnote{\url{https://claude.ai/}} large language model for such a list, which we show in App.~\ref{app:categories}.

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (CLIP-ViT-B/16). The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
  \label{tab:tti_base}
\end{table}

% \begin{figure*}[!h]
% \includegraphics[width=0.95\textwidth]{plots/tti_base.pdf}
% \caption{Image retrieval using the base CLIP model on four curated queries. \SM{Horizontal lines} \SM{Query not rotated} \SM{Same for next Figure.} \SM{Put text ``Base Model'' in consistent colour to legend.} \SM{Say top 4} \SM{Say that for science we care about the distributions of things.}}
% \label{fig:tti_base}
% \end{figure*}

% \begin{figure*}[!h]
% \includegraphics[width=0.95\textwidth]{plots/tti.pdf}
% \caption{Image retrieval using the fine-tuned CLIP model on four curated queries.}
% \label{fig:tti}
% \end{figure*}

@@

% \begin{figure*}[!h]
% \includegraphics[width=0.95\textwidth]{plots/itt.pdf}
% \caption{Text associations from a curated list most closely matching a given image query, for both the fine-tuned and base models. The `ground truth' LLM-summarized abstract is shown in the right column. \SM{Say ground truth summarized abstract in the column}\SM{Swap order of cols 2 and 3.}}
% \label{fig:itt}
% \end{figure*}

% Set line spacing to 0.5

\begin{table}[h!]
  \centering
  \renewcommand{\arraystretch}{0.1}
  \begin{tabular}{m{3cm} m{3.6cm} m{3.6cm} m{4cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_0.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_3.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (CLIP-ViT-B/16) and \textcolor{deepred}{summary fine-tuned} models. The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
  \label{tab:itt}
\end{table}

@@

The code, dataset, and fine-tuned models used in this work are available at \url{https://www.github.com/smsharma/HubbleCLIP}.

@@

% The regex-guided generation method ensures that the LLM's output always adheres to the specified regular expression pattern, without incurring the computational overhead of checking the validity of each possible next token at every generation step. This approach can be extended to more complex patterns, such as context-free grammars, by using augmented pushdown automata to track the generation state and valid next tokens.

% In our application, we use the regex-guided generation method to constrain LLM outputs to follow specific JSON schemas, which are automatically converted into regular expressions. This ensures that the generated summaries of astronomy observation proposals adhere to the desired format, enabling more reliable extraction of key information.
% The pre-computed index $\sigma$ allows for efficient determination of valid next tokens at each generation step, making the regex-guided method computationally feasible for large-scale applications.

@@

% \subsection{Generation of single-concept summaries}
% \label{app:singleconcept}

% The following prompt is used to generate a list of diverse single-concept summaries informed by a list of summarized abstracts:

% \begin{lstlisting}[language=Python]
% import outlines

% @outlines.prompt
% def prompt_fn(objects):
%     """<s>[INST] Please produce a list of around concepts characterizing prominent objects, phenomena, and science use cases of images observed by the Hubble Space Telescope.

% Here are some examples of objects:

% {{objects}}

% Follow these instructions exactly in your answer:
% - Do not output empty strings as elements.
% - Make sure that the list covers a diverse range of astronomical concepts, with items as different from each other as possible. 
% - Do not give specific names of objects, to make sure you span the widest possible range of concepts (e.g., "dwarf galaxy" is allowed, but NOT "Fornax", "Terzan 5", or  "NGC6440").
% - Do not return terms undescriptive of observations, e.g. "sloshing", "adiabatic", "interactions". Returning concrete physics objects, concepts, or phenomena.
% - Only output scientifically meaningful terms. E.g., NO "Cosmic Dance".
% - Do not duplicate entries. Do not reference any telescopes, observatories, or surveys.
% - Do not include units like "angular diameter distance", "parsec", or any other concepts that will not correlate with images of observations.
% - Use the above example list of objects only as inspiration to infer broad classes of objects.
% - Make sure each concept is succint, never more than 5 words.
% - Answer in JSON format.
% - The JSON should have the following keys {"galaxies", "stellar_physics", "exoplanets_planet_formation", "stellar_populations", "supermassive_black_holes", "solar_system", "integalactic_medium", "large_scale_structure"} reflecting rough observation categories.
% - Each category will have a list of objects and/or astronomical concepts.
% - Output up to 20 items and no more in each category
% [/INST]
% """
% \end{lstlisting}

% The following schema guides generation, intended to produce 100 concepts reflecting the science categories of successful proposals\footnote{\url{https://www.stsci.edu/contents/newsletters/2023-volume-40-issue-02/hubble-cycle-31-proposal-selection}}.
% %

% \begin{lstlisting}[language=Python]
% from pydantic import BaseModel, conlist

% class ScienceCategoriesHST(BaseModel):
%     galaxies: conlist(str, min_length=15, max_length=15)
%     stellar_physics: conlist(str, min_length=15, max_length=15)
%     exoplanets_planet_formation: conlist(str, min_length=15, max_length=15)
%     stellar_populations: conlist(str, min_length=10, max_length=10)
%     supermassive_black_holes: conlist(str, min_length=15, max_length=15)
%     solar_system: conlist(str, min_length=10, max_length=10)
%     integalactic_medium: conlist(str, min_length=10, max_length=10)
%     large_scale_structure: conlist(str, min_length=10, max_length=10)
%   \end{lstlisting}

% The output is then constrained to be one of the fixed number of concepts.

% \subsection{Assignment of abstracts to single-concept categories}
% \label{app:singleconceptassignments}

% Finally, the following prompt is used to assign concepts inferred in \ref{app:singleconcept} to each abstract:

% \begin{lstlisting}[language=Python]
% import outlines

% @outlines.prompt
% def prompt_fn(abs, cats):
%     """<s>[INST] The following is a successful proposal abstract for the Hubble Space Telescope: "{{abs}}"

% The following is a list of categories (astronomical concepts) that this abstract could correspond to.

% {{cats}}

% Please answer which of these listed concepts best describes this proposal, based on the objects and phenomena mentioned in the abstract.
% The concept should meaningfully be present in the abstract and the eventual observation.

% - For example, "The locations of supernovae {SNe} in the local stellar and gaseous environment in galaxies, as measured in high spatial resolution WFPC2 and ACS images, contain important clues to their progenitor stars." should return "supernova".
% - If the abstract centers calibration and/or instrumentation efforts, return calibration or instrumention".

% If no concept make sense, return "None". [/INST]
% """
% \end{lstlisting}

>>>>

\newcommand{\githubmaster}{\href{https://www.github.com/smsharma/PAPERCLIP-Hubble}{\faGithub}\xspace}

@@

\lstdefinestyle{mystyle}{ 
  backgroundcolor=\color{backcolour}, 
  commentstyle=\color{deepgreen},
  keywordstyle=\color{deepred}, 
  numberstyle=\tiny\color{codegray}, 
  stringstyle=\color{deepgreen},
  basicstyle=\ttfamily\footnotesize\linespread{1.1},
  breakatwhitespace=false, 
  breaklines=true, captionpos=b, 
  keepspaces=true, numbers=left, 
  numbersep=8pt, showspaces=false, 
  showstringspaces=false, showtabs=false, 
  frame=single, 
  framerule=0.2pt, 
  rulecolor=\color{codegray}, 
  tabsize=2, 
  aboveskip=1.5ex, 
  belowskip=1.5ex,
  xleftmargin=15pt, 
  xrightmargin=15pt, 
  extendedchars=true, 
  columns=flexible, 
  linewidth=\textwidth 
  }

@@

Machine learning (ML) is starting to have a significant impact in the sciences, with astrophysics being no exception.
%
ML methods have demonstrated promise at every stage of the research pipeline, from instrument design, to data acquisition, to its analysis \citep{huertas2022dawes}.
%
Until recently, most applications of ML within astrophysics have focused on augmenting traditional techniques in order to improve performance on specific tasks.
%
The {Foundation Model} paradigm, in contrast, seeks to develop generalist models which can be deployed to simultaneously tackle a wide range of tasks \citep{bommasani2021opportunities}.
%
These models are typically pre-trained on massive amounts of unlabeled data using self-supervised or weakly-supervised learning techniques, enabling them to learn powerful representations which can then be used downstream in different ways.
%
Foundation models can often benefit from additional training ({fine tuning}) using a relatively small amounts of domain-specific data in order to increase their usefulness when applied to specialized domains.

@@

In this paper, we describe \text{PAPERCLIP} (Proposal Abstracts Provide an Effective Representation for Contrastive Language-Image Pre-training\footnote{Technically, we fine tune rather than pre train, but ``PAPERCLIFT'' was rejected by the senior author of this paper.}), a method that connects astronomical image observations with natural language by leveraging the association between abstracts of successful observing proposals and images corresponding to downstream observations. 
%
Concretely, we showcase the method using observations imaged by the \hubble Space Telescope (HST).
%
We show that fine-tuning a pre-trained CLIP ~\citep[Contrastive Language-Image Pre-training; ][]{radford2021learning} image-text model on observation-abstract pairs results in meaningful joint representations through quantitative and qualitative evaluation tests.
%
Our method opens up the possibility of interacting with astronomical survey data using free-form natural language as an interface, which is a cornerstone of the success of the modern foundation model paradigm.
%

@@

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=0.99\textwidth]{plots/figure.pdf}
\caption{\changes{Overview of the PAPERCLIP method. (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts. (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query). The proposal abstract snippet here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}}
\label{fig:overview}
\end{figure*}

@@

@@

The guided generation process ensures that, in this case, the output of the LLM strictly conforms to the format of the following example:
\begin{center}
  \begin{jsoncode}
    \centering
          \color{black}\{
            \color{codegreen}'objects_and_phenomena'\color{black}: [\color{codegreen}'star forming galaxy', 'lensed galaxy'\color{black}], 
            \color{codegreen}'science_use_cases'\color{black}: [\color{codegreen}'measure lensing magnification'\color{black}]
          \color{black}\}
  \end{jsoncode}
  \end{center}
which is then used to construct the summarized caption by combining the two key elements.
%
Examples of raw abstract snippets and corresponding LLM-generated summaries are shown in Tab.~\ref{tab:datasetsumm}.
%
We train separate models using the raw abstracts and the LLM-generated summaries, and compare their performance on downstream tasks in Sec.~\ref{sec:results}.
%
We note that, even after summarization, the association signal is expected to be noisy, since parts of the summarized caption may not be directly descriptive of the observed images. The goal of the fine-tuning process is to leverage the signal contained in this noisy association.

@@

@@

The model fine-tuned with summarized abstracts, meanwhile, shows strikingly different behavior (Tab.~\ref{tab:tti}).
%
The \texttt{dwarf galaxy}-queried images correspond to proposals aiming to measure the kinematics of the stellar cores of dwarf galaxies.
%
Images of Jupiter are returned for the \texttt{Jupiter} query, with the planet clearly visible in all four images.
%
Supernova SN1987 itself can be seen in the three closest images for the \texttt{SN1987A} query with the fourth image being a supernova remnant.
%
Cluster-scale as well as galaxy-scale gravitational lenses are returned by the \texttt{strong lensing} query, with lensing patterns visible in the images.

@@

We can use images from the validation set as queries and retrieve the most relevant text chunks (e.g., objects and use cases) from a curated list as described in Sec.~\ref{sec:eval}.
%
We show the result of image-to-text retrieval in Tab.~\ref{tab:itt}, for the base (second column) as well as summary fine-tuned (third column) models, using four observations (left-most column) from the validation set.
%

@@

\begin{table}[h!]
  \centering
  \begin{tabular}{m{2.7cm} p{2.9cm} p{2.9cm} p{2.9cm} p{2.9cm}}
      \toprule
      \centering \bfseries Query & \multicolumn{4}{c}{\bfseries{Top-4 most similar images using \textcolor{deeppurple}{base CLIP model}}} \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_1.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_0.pdf} \\ \input{\thedatafolder/propid_tti_base_1_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_1.pdf} \\ \input{\thedatafolder/propid_tti_base_1_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_2.pdf} \\ \input{\thedatafolder/propid_tti_base_1_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_1_3.pdf} \\ \input{\thedatafolder/propid_tti_base_1_3.txt}  \tabularnewline
      \midrule
       \texttt{\input{\thedatafolder/query_tti_base_0.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_0.pdf} \\ \input{\thedatafolder/propid_tti_base_0_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_1.pdf} \\ \input{\thedatafolder/propid_tti_base_0_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_2.pdf} \\ \input{\thedatafolder/propid_tti_base_0_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_0_3.pdf} \\ \input{\thedatafolder/propid_tti_base_0_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_2.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_0.pdf} \\ \input{\thedatafolder/propid_tti_base_2_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_1.pdf} \\ \input{\thedatafolder/propid_tti_base_2_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_2.pdf} \\ \input{\thedatafolder/propid_tti_base_2_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_2_3.pdf} \\ \input{\thedatafolder/propid_tti_base_2_3.txt}  \tabularnewline
      \midrule
      \texttt{\input{\thedatafolder/query_tti_base_3.txt}} \vspace{20mm} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_0.pdf} \\ \input{\thedatafolder/propid_tti_base_3_0.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_1.pdf} \\ \input{\thedatafolder/propid_tti_base_3_1.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_2.pdf} \\ \input{\thedatafolder/propid_tti_base_3_2.txt} & \centering \includegraphics[width=0.18\textwidth]{\thedatafolder/img_tti_base_3_3.pdf} \\ \input{\thedatafolder/propid_tti_base_3_3.txt}  \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{For four text queries (left-most column), the four most similar images from the validation dataset by cosine similarity when using the \textbf{\textcolor{deeppurple}{base CLIP model}} (\texttt{CLIP-ViT-B/16}). The proposal ID associated with each image is given below the image and contains a hyperlink to the MAST page corresponding to the proposal.}
  \label{tab:tti_base}
\end{table}

@@

\begin{table}[h!]
  \centering
  \renewcommand{\arraystretch}{0.1}
  \begin{tabular}{m{3cm} m{3.6cm} m{3.6cm} m{4cm}}
      \toprule
      \centering \bfseries \hubble image & \centering \textbf{Top-4 text} \\ \small\textbf{\textcolor{deeppurple}{(base)}} & \centering  \textbf{Top-4 text} \\ \small\textbf{\textcolor{deepred}{(summary fine-tuned)}} & \centering \textbf{Summarized abstract} \\ (objects; `ground truth') \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_0.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_0.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_0.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_0.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_1.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_1.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_1.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_1.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_2.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_2.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_2.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_2.txt}} \tabularnewline
      \midrule
      \centering \includegraphics[width=0.15\textwidth]{\thedatafolder/img_itt_3.pdf} & \centering \scriptsize \verbatiminput{\thedatafolder/sci_itt_base_3.txt} & \centering  \scriptsize \verbatiminput{\thedatafolder/sci_itt_3.txt} &  {\scriptsize \input{\thedatafolder/abs_itt_3.txt}} \tabularnewline
      \bottomrule
  \end{tabular}
  \caption{Text snippets from a curated list most closely matching a given image query (left-most column) by cosine similarity of respective embeddings, shown for the \textcolor{deeppurple}{base} (\texttt{CLIP-ViT-B/16}) and \textcolor{deepred}{summary fine-tuned} models. The `ground truth' LLM-summarized abstract (only objects/phenomena) is shown in the right-most column.}
  \label{tab:itt}
\end{table}

@@

The code, dataset, and fine-tuned models used in this work are available at \url{https://www.github.com/smsharma/PAPERCLIP-Hubble}.

@@

@@
